A comparison of classification algorithms applied to the Adult Database

### 2009-12-09 report/final.tex
report adtree

@@

>>>>

\subsection{Alternating Decision Trees}

Alternating Decision Trees are related to a concept named \textit{boosting}.
Boosting is based on the observation that finding many rough rules of thumb can be a lot easier 
than finding a single, highly accurate prediction rule. To apply the boosting approach,
we start with a method or algorithm for finding the rough rules of thumb. 
The boosting algorithm calls this “weak” or “base” learning algorithm repeatedly, 
each time feeding it a different subset of the training examples. Each time 
it is called, the base learning algorithm generates a new weak prediction rule, and 
after many rounds, the boosting algorithm must combine these weak rules into a 
single prediction rule that, hopefully, will be much more accurate than any one of 
the weak rules. \cite{boosting}

The ADTree's Weka algorithm can be parametrized by the number of boosting iterations
\textbf{$<$-B integer$>$} and by the way the algorithm will expand the nodes,
\textbf{$<$-E [-3,-2,-1,$>=$0]$>$},
-3 for \textit{all}, -2 for \textit{weight} based expanding,
-1 for \textit{z\_pure} and a number
equal or greater than zero as a seed for random walks.

We experimented the algorithm using various combinations of those two parameters.

\begin{table}[ht]
  \begin{center}
  \begin{tabular}{ | l | c | c | c | c | c |}
    \hline
    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} & \textbf{Nodes} & \textbf{Leaves} \\ \hline
    -E -3 -B 2 & 21.80 & 21.44 & 0m24.641s & 7 & 5 \\ \hline
    -E -3 -B 5 & 18.20 & 20.99 & 0m24.571s & 16 & 11 \\ \hline
    -E -3 -B 10 & 16.13 & 20.40 & 0m26.539s & 31 & 21 \\ \hline
    -E -3 -B 15 & 15.91 & 20.19 & 0m28.565s & 46 & 31 \\ \hline
    -E -3 -B 20 & 15.25 & 20.39 & 0m32.336s & 61 & 41 \\ \hline
    -E -3 -B 25 & 15.08 & 20.37 & 0m37.223s & 76 & 51 \\ \hline
    -E -3 -B 30 & 14.88 & 20.25 & 0m41.062s & 91 & 61 \\ \hline
    -E -3 -B 35 & 14.78 & 20.01 & 0m45.760s & 106 & 71 \\ \hline
    -E -3 -B 40 & 14.68 & 20.58 & 0m52.976s & 121 & 81 \\ \hline

    -E -2 -B 2 & 21.80 & 21.44 & 0m25.219s & 7 & 5 \\ \hline
    -E -2 -B 5 & 18.20 & 20.99 & 0m24.445s & 16 & 11 \\ \hline
    -E -2 -B 10 & 16.34 & 20.71 & 0m25.390s & 31 & 21 \\ \hline
    -E -2 -B 15 & 15.80 & 20.42 & 0m26.982s & 46 & 31 \\ \hline
    -E -2 -B 20 & 15.13 & 19.80 & 0m29.050s & 58 & 39 \\ \hline
    -E -2 -B 25 & 14.95 & 20.25 & 0m28.956s & 73 & 49 \\ \hline
    -E -2 -B 30 & 14.93 & 20.07 & 0m31.118s & 88 & 59 \\ \hline
    -E -2 -B 35 & 14.90 & 20.20 & 0m32.565s & 100 & 67 \\ \hline
    -E -2 -B 40 & 14.65 & 20.73 & 0m34.003s & 115 & 77 \\ \hline

    -E -1 -B 2 & 21.80 & 21.436 & 0m24.149s & 7 & 5 \\ \hline
    -E -1 -B 5 & 18.20 & 20.99 & 0m24.061s & 16 & 11 \\ \hline
    -E -1 -B 10 & 16.34 & 20.71 & 0m24.967s & 31 & 21 \\ \hline
    -E -1 -B 15 & 15.80 & 20.42 & 0m27.095s & 46 & 31 \\ \hline
    -E -1 -B 20 & 15.13 & 19.80 & 0m28.813s & 58 & 38 \\ \hline
    -E -1 -B 25 & 14.95 & 20.25 & 0m29.205s & 73 & 49 \\ \hline
    -E -1 -B 30 & 14.93 & 20.07 & 0m30.569s & 88 & 59 \\ \hline
    -E -1 -B 35 & 14.93 & 20.20 & 0m32.708s & 100 & 67 \\ \hline
    -E -1 -B 40 & 14.65 & 20.73 & 0m33.205s & 115 & 77 \\ \hline

    -E 0 -B 2 & 21.80 & 21.44 & 0m23.717s & 7 & 4 \\ \hline
    -E 0 -B 5 & 18.21 & 20.99 & 0m24.619s & 16 & 11 \\ \hline
    -E 0 -B 10 & 16.37 & 20.71 & 0m24.512s & 31 & 21 \\ \hline
    -E 0 -B 15 & 15.81 & 20.42 & 0m25.240s & 46 & 31 \\ \hline
    -E 0 -B 20 & 14.94 & 20.07 & 0m25.589s & 61 & 41 \\ \hline
    -E 0 -B 25 & 15.04 & 20.15 & 0m25.868s & 76 & 51 \\ \hline
    -E 0 -B 30 & 15.01 & 20.50 & 0m26.492s & 91 & 61 \\ \hline
    -E 0 -B 35 & 14.96 & 20.53 & 0m26.294s & 106 & 71 \\ \hline
    -E 0 -B 40 & 14.90 & 20.54 & 0m27.248s & 118 & 79 \\ \hline

  \end{tabular}
  \caption{Results for ADTree.}
  \label{tbl:results_adtree}
  \end{center}
\end{table}

From Table \ref{tbl:results_adtree} we can see that the number of boosting iterations affects
both the execution time (directly proportional) and the error rate.
The best error rate is for the parameters \textbf{-E -2 -B 20} giving $\approx$19.80\% for the test set.

Another good observation is that this algorithm can build small trees that give satisfactory results, both for
the train set and for the test set.

@@

\bibitem{boosting}
Robert E. Schapire, The Boosting Approach to Machine Learning: An Overview, 2001.
